\documentclass{egpubl}
\usepackage{eurovis2021}

% --- for  Annual CONFERENCE
% \ConferenceSubmission   % uncomment for Conference submission
% \ConferencePaper        % uncomment for (final) Conference Paper
% \STAR                   % uncomment for STAR contribution
% \Tutorial               % uncomment for Tutorial contribution
% \ShortPresentation      % uncomment for (final) Short Conference Presentation
% \Areas                  % uncomment for Areas contribution
% \MedicalPrize           % uncomment for Medical Prize contribution
% \Education              % uncomment for Education contribution
% \Poster                 % uncomment for Poster contribution
% \DC                     % uncomment for Doctoral Consortium
%
% --- for  CGF Journal
% \JournalSubmission    % uncomment for submission to Computer Graphics Forum
% \JournalPaper         % uncomment for final version of Journal Paper
%
% --- for  CGF Journal: special issue
% \SpecialIssueSubmission    % uncomment for submission to , special issue
\SpecialIssuePaper         % uncomment for final version of Computer Graphics Forum, special issue
%                          % EuroVis, SGP, Rendering, PG
% --- for  EG Workshop Proceedings
% \WsSubmission      % uncomment for submission to EG Workshop
% \WsPaper           % uncomment for final version of EG Workshop contribution
% \WsSubmissionJoint % for joint events, for example ICAT-EGVE
% \WsPaperJoint      % for joint events, for example ICAT-EGVE
% \Expressive        % for SBIM, CAe, NPAR
% \DigitalHeritagePaper
% \PaperL2P          % for events EG only asks for License to Publish

% --- for EuroVis 
% for full papers use \SpecialIssuePaper
% \STAREurovis   % for EuroVis additional material 
% \EuroVisPoster % for EuroVis additional material 
% \EuroVisShort  % for EuroVis additional material

% !! *please* don't change anything above
% !! unless you REALLY know what you are doing
% ------------------------------------------------------------------------

\usepackage[T1]{fontenc}
\usepackage{dfadobe}


% \usepackage{cite}  % comment out for biblatex with backend=biber
% ---------------------------
\biberVersion
\BibtexOrBiblatex
\usepackage[backend=biber,bibstyle=EG,citestyle=alphabetic,backref=true,url=false]{biblatex}
%\addbibresource{egbibsample.bib}

\renewbibmacro*{doi+eprint+url}{%
    \printfield{doi}%
    \newunit\newblock%
    \iftoggle{bbx:eprint}{%
        \usebibmacro{eprint}%
    }{}%
    \newunit\newblock%
    \iffieldundef{doi}{%
        \usebibmacro{url+urldate}}%
        {}%
    }

\addbibresource{GeoVis.bib}

\newcommand{\citetitlea}[1]{\citefield{#1}{title}}

% ---------------------------  
\electronicVersion
\PrintedOrElectronic
% for including postscript figures
% mind: package option 'draft' will replace PS figure by a filename within a frame
\ifpdf \usepackage[pdftex]{graphicx} \pdfcompresslevel=9
\else \usepackage[dvips]{graphicx} \fi

\usepackage{egweblnk}
% end of prologue

\usepackage{float,hhline,hyperref,array,tabulary,tabularx,etoolbox,multirow}

\usepackage{dblfloatfix}

\usepackage[table, dvipsnames]{xcolor}

\usepackage{multicol}

\include{preamble/preamble.basic}
\include{preamble/preamble.main}
\include{preamble/preamble.math}
\include{preamble/preamble.draft}

\author[Q. Wang \& R. S. Laramee]
{\parbox{\textwidth}{\centering Q. Wang
  and R. S. Laramee
    }
    \\
\parbox{\textwidth}{\centering School of Computer Science, University of Nottingham, UK}
}

\title{Cartograms with Dynamic Features}

\begin{document}

\pagestyle{plain}

\maketitle

\begin{abstract}
Cartograms are hybrid representations of geographical and abstract data based on a value-by-area mapping technique. Based on the Dorling cartogram, the Demers cartogram uses squares instead of circles to represent regions, enabling the comparison of regions in a more intuitive way with a higher screen space utilization. One of the disadvantages of the Dorling cartogram and its variants is that regions may be displaced from their original positions, reducing the legibility, readability, and geographical accuracy. To address that, we propose a novel layout algorithm that integrates dynamic topological features such as rivers into Demers cartograms. We evaluate the proposed layout algorithm through a user study with an application to an Electronic Health Records (EHR) dataset. The results demonstrate that the proposed layout algorithm is able to preserve some of the original properties of the cartogram, while improving the legibility, readability, and geographical accuracy of the cartogram. 
\end{abstract}

\input{section/introduction}

\input{section/related}

\input{section/data}

\input{section/main}

\section{Evaluation}

\section{Limitations and Future Work}

\section{Conclusions}

\section{Acknowledgments}
This work is funded by the grant EP/S010238/2 from the Engineering and Physical Sciences Research Council (EPSRC). EPSRC is a British Research Council that provides government funding for grants to undertake research and postgraduate degrees in engineering and the physical sciences.


\clearpage
% reset \section command for section title "References"
\let\section=\origsection
\printbibliography

\cleardoublepage

\input{section/appendix}

\end{document}
